Generation of TypeScript Declaration Files from JavaScript Code

### 2020-08-07 programming-2021/main.tex
section 2, 3 examples

\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{positioning}

@@

\lstdefinelanguage{JavaScript}{
  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, undefined},
  keywordstyle=\color{black}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this},
  ndkeywordstyle=\color{black}\bfseries,
  identifierstyle=\color{darkgray},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{lightgray}\slshape,
  stringstyle=\color{darkgray}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

\lstdefinelanguage{TypeScript}{
  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, undefined},
  keywordstyle=\color{black}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this, declare, constructor, namespace, interface},
  ndkeywordstyle=\color{black}\bfseries,
  identifierstyle=\color{darkgray},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{lightgray}\slshape,
  stringstyle=\color{darkgray}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

@@

\section{Introduction}
\label{sec:introduction}
JavaScript is the most popular language for writing web
applications \cite{github-statistics}. It is also gaining popularity
for back-end applications running in \NodeJS{}, a JavaScript-based
server-side platform. JavaScript is appealing to developers because
its forgiving dynamic typing enables 
them to create simple pieces of code very quickly and proceed on a
trial-and-error basis.

JavaScript was never intended to be more than a
scripting language and thus lacks features for maintaining and evolving large
codebases. However, JavaScript is increasingly used for creating large and complex
applications. 
Mistakes such as mistyped property
names and misunderstood or unexpected type coercions cause developers
to spend a significant amount of time in debugging. There is ample
evidence for such mishaps. For example, a 
JavaScript code blog\footnote{\url{https://wtfjs.com}} collects experiences
from developers facing unexpected situations while programming in
JavaScript. \coderef{code:introduction-javascript-wtfs} exposes some
of these unintuitive JavaScript behaviors.

\begin{lstlisting}[
    caption={\textbf{Unintuitive JavaScript behavior} - falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion},
    language=JavaScript,
	label=code:introduction-javascript-wtfs,
    float,
    abovecaptionskip=-\medskipamount
]
"0" == false; // true
true == "1.00"; // true
false == "    \n\r\t     "; // true
false == []; // true
0 == []; // true
null == undefined; // true
[1] + 1; //'11'
[2] == "2"; // true
null + undefined + [1, 2, 3] // 'NaN1,2,3'
"hello world".lenth + 1 // NaN | note 'lenth' instead of 'length'
[1, 15, 20, 100].sort() // [ 1, 100, 15, 20 ]
typeof null; // object
null instanceof Object; // false
"01" < "00100"; // false
\end{lstlisting}
% pjt: why unexpected?

The cognitive load produced by these behaviors is mitigated in languages that use
build tools based on type information. This insight motivated the
creation of TypeScript, a superset of JavaScript with expressive type
annotations \cite{typescript}. It has become a widely used alternative
among JavaScript developers, because it incorporates features that are
helpful for developing and maintaining large applications
\cite{DBLP:conf/icse/GaoBB17}. TypeScript enables the early detection
of several kinds of run-time errors and the integration of code intelligence
tooling like auto-completion in an IDE.

Despite the advantages, it is unrealistic to expect the world to
switch to TypeScript in a day. Therefore, 
existing JavaScript libraries can be used in a TypeScript project by
adding a declaration file that contains a description of the library's
API in terms of types. Lots of declaration files for popular
JavaScript libraries have been created in a community effort in the
DefinitelyTyped repository~\cite{definitely-typed-repository}.
At the time of writing this repository contains declaration files for
more than 6000 JavaScript libraries. Unfortunately, most declaration
files in this repository have been manually created and maintained,
which is error prone and time consuming. As TypeScript takes a
declaration file at face value, discrepancies are not detected at
compile time and the resulting inaccurate code-intelligence features
are confusing for the
programmer. Moreover, TypeScript does not perform any run-time 
checking of types, either, so that a discrepancy  between the declaration file
and its corresponding JavaScript library can lead to unexpected 
behavior and crashes. Such behavior can lead to developer
frustration, longish debugging sessions, and decreasing confidence in
the tool chain.

We aim to improve on this situation by automatically generating
sound TypeScript declaration files. To this end, we rely on the examples provided
by programmers as part of the documentation of their library. If these
examples are sufficiently rich, then our toolchain can generate the
desired declaration files from them. This approach takes advantage of
best practices in the dynamic languages community which favors
examples over writing out type signatues.

\begin{description}
\item[PJT: How about a short example at this point?] 
\item[PJT: maybe later] Some previous work tackled the problem of improving
  the quality of declaration files. Feldthaus et al
  \cite{DBLP:conf/oopsla/FeldthausM14} search automatically for
  mismatches between a declaration file and implementation
  code. TSTest \cite{DBLP:journals/pacmpl/KristensenM17} adapted
  feedback-directed random testing to detect discrepancies between a
  declaration file and a JavaScript library. Tools like TSInfer and
  TSEvolve~\cite{DBLP:conf/fase/KristensenM17} are designed for
  assisting the creation of new declaration files and supporting the
  evolution the declaration file when the corresponding JavaScript
  library gets modified, respectively. They rely on an existing static
  analyzer for JavaScript \cite{DBLP:conf/sas/JensenMT09}.  TypeScript
  itself developed \texttt{dts-gen}, a tool that generates a
  declaration file that is meant to be used as a \emph{starting point
    for writing a high-quality declaration file} \cite{dts-gen}.
\end{description}
In this work we present the tool \texttt{dts-generate}
as a first step to explore the possibilities for
generating useful declaration files without the heavy lifting of
static analysis. \texttt{dts-generate} comes with a framework that
supports the generation of declaration files for an existing
JavaScript library published to the \NPM{} registry. The tool gathers
data flow and type information at run-time to generate a declaration
file based on that information.

The novelty of our tool is twofold:
\begin{enumerate}
\item 
  We do not rely on static analysis, which is hard to implement
  soundly and precisely and which is prone to maintenance problems
  when keeping up with JavaScript's yearly language updates.
\item
  Instead we extract example code from the programmer's library
  documentation and rely on dynamic analysis to extract typed usage
  patterns for the library from the example runs.
\end{enumerate}
% The architecture supports the future incorporation of a Symbolic
% Execution Engine that expands the initial code base using the
% signatures in the declaration file. The iterative process of exploring
% new execution paths will refine the generated declaration file in each
% iteration.

@@

The contributions of this paper are as follows.
\begin{itemize}
% \item We introduce an architecture that supports generating TypeScript
%   Declaration Files for a given JavaScript Library using run-time
%   information. The architecture supports the future incorporation of a
%   Symbolic Execution Engine that expands the initial code base using
%   the signatures in the declaration file. The iterative process of
%   exploring new execution paths will refine the generated declaration
%   file in each iteration.
\item A framework that extracts code examples from the
  documentation of an \NPM{} package and collects run-time type
  information from running these examples.

\item The tool \texttt{dts-generate}, a command line
  application that generates a valid TypeScript declaration file for a
  specific \NPM{} package using run-time information.

@@

\section{Motivating Example}
\label{sec:motivating-example}
The \NPM{} package \texttt{glob-to-regexp} is a simple JavaScript library, which turns a glob expression into a regular expression\footnote{https://www.npmjs.com/package/glob-to-regexp}. It
has about 6,500,000 weekly downloads and 181 \NPM{} packages depend on it. If
a developer creates or extends TypeScript code that depends on the
\texttt{glob-to-regexp} library, the TypeScript
compiler and IDE requires a declaration file for that library to
perform static checking and code completion, respectively. We 
use \texttt{dts-generate} to automatically generate a TypeScript
declaration file for \texttt{glob-to-regexp}. The tool downloads the \NPM{} 
package, runs the examples extracted from its documentation, gathers
run-time information, and generates a TypeScript declaration
file. The result is 
shown in \figref{fig:motivating-example-glob-to-regexp-vscode} and it is
ready for use in a TypeScript project. For example,  Visual Studio's
code completion runs properly with it
(\figref{fig:motivating-example-glob-to-regexp-vscode}). If the
\texttt{glob-to-regexp} package gets modified in the future, a new declaration
file can be generated automatically using
\texttt{dts-generate}. Our comparator tool can compare the new file
for incompatibilities with the previous declaration file.

\begin{figure}[tp]
  \centering
  \begin{subfigure}{0.70\linewidth}
    \begin{lstlisting}[language=TypeScript]
export = GlobToRegexp;

declare function GlobToRegexp(glob: string, opts: undefined): RegExp;
declare function GlobToRegexp(glob: string, opts: GlobToRegexp.I__opts): RegExp;
declare namespace GlobToRegexp {
  export interface I__opts {
    'extended': undefined | boolean;
    'globstar': undefined | boolean;
    'flags': undefined | string;
  }

}
    \end{lstlisting}
  \end{subfigure}

  \hspace{10.\textwidth}

  \begin{subfigure}{1.\linewidth}
    \centering
    \includegraphics[width=0.8\linewidth]{motivating-example-glob-to-regexp-vscode.png}
  \end{subfigure}

  \caption{\textbf{Declaration file for \texttt{glob-to-regexp} generated with \texttt{dts-generate}} - Interfaces is correctly detected. Optional parameters are explicitly declared as \texttt{undefined}. Declaration file can be correctly used in Visual Studio Code\footnote{https://code.visualstudio.com}.}
  \label{fig:motivating-example-glob-to-regexp-vscode}
\end{figure}

After filling in some background information on declaration files,
Section~\ref{sec:gener-typescr-decl} examines each step in the
generation process in detail and refer back to this example for
concreteness.

@@

The declaration file shown in
\figref{fig:motivating-example-glob-to-regexp-vscode} describes a
package with a single exported function. The first parameter is of type \texttt{string} and the second one is an optional interface. TypeScript namespaces are used for organizing types declared in a declaration file and avoiding name collisions with other types \cite{typescript-namespaces}. For example, the interface \texttt{I\_\_opts} declared in \figref{fig:motivating-example-glob-to-regexp-vscode} is used as \lstinline[language=TypeScript]{globToRegexp.I__opts}, since it is declared within a namespace.

This file is an instance of one of the standard templates for writing
declaration
files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/templates.html}:
\textbf{module}, \textbf{module-class} and
\textbf{module-function}. 
Each template corresponds to a different way of describing the exports
of a JavaScript library. Choosing the template depends on the
particular organization of the underlying JavaScript library:
\begin{description}
\item[module] several exported functions,
\item[module-class] a class-like structure,
\item[module-function] exactly one exported functions.
\end{description}

@@

TypeScript provides a guide on how to write high quality declaration files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/by-example.html}. They explain the main concepts through examples. We extracted one example for each template and generated used \texttt{dts-generate} to generate a corresponding declaration file, as presented in \figref{fig:typescript-templates-by-example}. The \texttt{glob-to-regexp} library is an instance of a \textbf{module-function}. It is worth mentioning that templates are a way of describing how a JavaScript library is intended to be used. Although highly unlikely, it is possible to create a library that can be represented with a different template depending on how it is used, as shown in \figref{fig:multi-purpose-module-example}. Run-time analysis proves useful in this case, since it is not sufficient to statically analyze the exported module.

\begin{figure}[tp]
  \centering
  \begin{subfigure}{0.48\linewidth}
    \begin{lstlisting}[language=JavaScript]
var greet = require("./greet-settings-module");

@@

greet({
  greeting: "hello world",
  color: "#00ff00"
});
    \end{lstlisting}
    \caption{TypeScript example for \texttt{module-function} template}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.48\linewidth}
    \begin{lstlisting}[language=TypeScript]
export = GreetSettingsModule;

declare function GreetSettingsModule(settings: GreetSettingsModule.I__settings): void;
declare namespace GreetSettingsModule {
  export interface I__settings {
    'greeting': string;
    'duration': number | undefined;
    'color': undefined | string;
  }

@@

  \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=JavaScript]
var myLib = require("./greet-module");

@@

var goodbye = myLib.makeGoodBye();
console.log("The computed goodbye is: " + goodbye);    
      \end{lstlisting}
      \caption{Example for \texttt{module}}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export function makeGreeting(str: string): string;
export function makeGoodBye(): string;        
      \end{lstlisting}
      \caption{Declaration file for \texttt{module} template}
    \end{subfigure}

    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=JavaScript]
var Greeter = require("./greet-classes-module.js");

var myGreeter = new Greeter("hello, world");
myGreeter.greeting = "howdy";
myGreeter.showGreeting();
      \end{lstlisting}
      \caption{Example for \texttt{module-class}}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export = Greeter;

@@

  \caption{Results for \textbf{module-function} template}
  \label{fig:typescript-templates-by-example}
\end{figure}

\begin{figure}[tp]
  \centering
  \begin{subfigure}{0.80\linewidth}
      \begin{lstlisting}[language=JavaScript]
var multiPurposeModule = require('./multi-purpose-module');

// module-function
console.log(multiPurposeModule("John", "Doe"));
// John Doe

// module-class
var m = new multiPurposeModule("Jane", "Doe");
console.log(m.firstName + " - " + m.lastName);
// Jane - Doe

console.log(m.sayHello());
// Hello, my name is Jane Doe

// module
console.log(multiPurposeModule.anotherFunction());
// I am another function! 
      \end{lstlisting}
      \caption{index.js}
    \end{subfigure}

    \begin{subfigure}{0.80\linewidth}
      \begin{lstlisting}[language=TypeScript]
var multiPurposeModule = function(firstName, lastName) {
  this.firstName = firstName;
  this.lastName = lastName;

  return firstName + " " + lastName;
};

multiPurposeModule.prototype.sayHello = function() {
  return "Hello, my name is " + this.firstName + " " + this.lastName;
};

multiPurposeModule.anotherFunction = function() {
  return "I am another function!";
};

module.exports = multiPurposeModule;   
      \end{lstlisting}
      \caption{multi-purpose-module/index.js}
    \end{subfigure}
  \caption{Example module that implements all three templates simultaneously}
  \label{fig:multi-purpose-module-example}
\end{figure}

@@

\begin{figure}[tp]
  \centering
  \begin{subfigure}{0.48\linewidth}
    \begin{lstlisting}[language=JavaScript]
var multiPurposeModule = require('./multi-purpose-module');

multiPurposeModule("John", "Doe");
// John Doe
    \end{lstlisting}
    \caption{Example module \texttt{multi-purpose-module} used as \texttt{module-function}}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.48\linewidth}
    \begin{lstlisting}[language=TypeScript]
export = MultiPurposeModule;

declare function MultiPurposeModule(firstName: string, lastName: string): string;
    \end{lstlisting}
    \caption{Generated declaration file with \texttt{module-function} template}
  \end{subfigure}

  \begin{subfigure}{0.48\linewidth}
    \begin{lstlisting}[language=JavaScript]
var multiPurposeModule = require('./multi-purpose-module');

var m = new multiPurposeModule("Jane", "Doe");
var fullName = m.firstName + " - " + m.lastName;
// Jane - Doe

m.sayHello();
// Hello, my name is Jane Doe
    \end{lstlisting}
    \caption{Example module \texttt{multi-purpose-module} used as \texttt{module-class}}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export = MultiPurposeModule;

declare class MultiPurposeModule {
  constructor(firstName: string, lastName: string);
  sayHello(): string;
}

declare namespace MultiPurposeModule {
}
      \end{lstlisting}
      \caption{Generated declaration file with \texttt{module-class} template}
    \end{subfigure}

    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=JavaScript]
var multiPurposeModule = require('./multi-purpose-module');

multiPurposeModule.anotherFunction();
// I am another function!
      \end{lstlisting}
      \caption{Example module \texttt{multi-purpose-module} used as \texttt{module}}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript]
export function anotherFunction(): string;
      \end{lstlisting}
      \caption{Generated declaration file with \texttt{module} template}
    \end{subfigure}

>>>>

\usepackage{cleveref}
\usepackage{pgfplots}
\usepackage{tikz}
\usetikzlibrary{positioning}

@@

\lstdefinelanguage{JavaScript}{
  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, undefined,const},
  keywordstyle=\color{black}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this},
  ndkeywordstyle=\color{black}\bfseries,
  identifierstyle=\color{darkgray},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{lightgray}\slshape,
  stringstyle=\color{darkgray}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

\lstdefinelanguage{TypeScript}{
  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, undefined,const},
  keywordstyle=\color{black}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this, declare, constructor, namespace, interface},
  ndkeywordstyle=\color{black}\bfseries,
  identifierstyle=\color{darkgray},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{lightgray}\slshape,
  stringstyle=\color{darkgray}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

@@

\section{Introduction}
\label{sec:introduction}
JavaScript is the most popular language for writing web
applications \cite{github-statistics}. It is also increasingly used
for back-end applications running in \NodeJS{}, a JavaScript-based
server-side platform. JavaScript is appealing to developers because
its forgiving dynamic typing enables 
them to create simple pieces of code very quickly and proceed on a
trial-and-error basis.

JavaScript was never intended to be more than a
scripting language and thus lacks features for maintaining and evolving large
codebases. However, nowadays developers create large and complex
applications in JavaScript. 
Mistakes such as mistyped property
names and misunderstood or unexpected type coercions cause developers
to spend a significant amount of time in debugging. There is ample
evidence for such mishaps. For example, a 
JavaScript code blog\footnote{\url{https://wtfjs.com}} collects experiences
from developers facing unexpected situations while programming in
JavaScript. \coderef{code:introduction-javascript-wtfs} exposes some
of these unintuitive JavaScript behaviors.

\begin{lstlisting}[
    caption={\textbf{Unintuitive JavaScript behavior} Falsy values, \lstinline!typeof!, \lstinline!null! and \lstinline!undefined! operators and type coercion},
    language=JavaScript,
	label=code:introduction-javascript-wtfs,
    float=bp,
    abovecaptionskip=-\medskipamount
]
"0" == false; // true
true == "1.00"; // true
false == "    \n\r\t     "; // true
false == []; // true
0 == []; // true
null == undefined; // true
[1] + 1; //'11'
[2] == "2"; // true
null + undefined + [1, 2, 3] // 'NaN1,2,3'
"hello world".lenth + 1 // NaN | note 'lenth' instead of 'length'
[1, 15, 20, 100].sort() // [ 1, 100, 15, 20 ]
typeof null; // object
null instanceof Object; // false
"01" < "00100"; // false
\end{lstlisting}
% pjt: why unexpected?

The cognitive load produced by these behaviors is mitigated in languages that use
build tools based on type information. This insight motivated the
creation of TypeScript, a superset of JavaScript with expressive type
annotations \cite{typescript}. It has become a widely used alternative
among JavaScript developers, because it incorporates features that are
helpful for developing and maintaining large applications
\cite{DBLP:conf/icse/GaoBB17}. TypeScript enables the early detection
of several kinds of run-time errors and the integration of code intelligence
tools like autocompletion in an IDE.

Despite the advantages, it is unrealistic to expect the world to
switch to TypeScript in a day. Therefore, 
existing JavaScript libraries can be used in a TypeScript project by
adding a declaration file that contains a description of the library's
API in terms of types. 
The DefinitelyTyped repository~\cite{definitely-typed-repository} has
been created as a community effort to collect declaration files for
popular JavaScript libraries. At the time of writing it contains
declaration files for more than 6000 libraries.

Unfortunately, creation and maintenance for most declaration
files in DefinitelyTyped is conducted manually,
which is error prone and time consuming. As TypeScript takes a
declaration file at face value, discrepancies are not detected at
compile time and the resulting inaccuracies in autocompletion, say,
are confusing for the programmer.
As TypeScript does not perform any run-time 
checking of types, either,  discrepancies  between the declaration file
and its corresponding JavaScript library can lead to unexpected 
behavior and crashes. Such an experience can lead to developer
frustration, longish debugging sessions, and decreasing confidence in
the tool chain.

\subsection{Approach}
\label{sec:approach}

We aim to improve on this situation by providing a tool that generates
sound TypeScript declaration files automatically. To do so, we rely on the examples provided
as part of the documentation of a library. If these
examples are sufficiently rich, then our toolchain can generate
high-quality declaration files from them. This approach takes
advantage of 
best practices in the dynamic languages community which favors
examples and tests over writing out type signatures.

As an example, consider the NPM module
\texttt{abs} that
``computes the absolute path of an input''. Its documentation consists
of the following three examples\footnote{\url{https://www.npmjs.com/package/abs}}.
\begin{lstlisting}[language=JavaScript,numbers=none]
const abs = require("abs");

console.log(abs("/foo"));
// => "/foo"

console.log(abs("foo"));
// => "/path/to/where/you/are/foo"

console.log(abs("~/foo"));
// => "/home/username/foo"
\end{lstlisting}
From the examples, our tool generates a TypeScript declaration file,
which is equivalent to the declaration file provided in
DefinitelyTyped\footnote{\url{https://github.com/DefinitelyTyped/DefinitelyTyped/blob/master/types/abs/index.d.ts}}: 
\begin{lstlisting}[language=TypeScript,numbers=none]
export = Abs;
declare function Abs(input: string): string;
\end{lstlisting}
% \begin{description}
% \item[PJT: maybe later] Some previous work tackled the problem of improving
%   the quality of declaration files. Feldthaus et al
%   \cite{DBLP:conf/oopsla/FeldthausM14} search automatically for
%   mismatches between a declaration file and implementation
%   code. TSTest \cite{DBLP:journals/pacmpl/KristensenM17} adapted
%   feedback-directed random testing to detect discrepancies between a
%   declaration file and a JavaScript library. Tools like TSInfer and
%   TSEvolve~\cite{DBLP:conf/fase/KristensenM17} are designed for
%   assisting the creation of new declaration files and supporting the
%   evolution the declaration file when the corresponding JavaScript
%   library gets modified, respectively. They rely on an existing static
%   analyzer for JavaScript \cite{DBLP:conf/sas/JensenMT09}.  TypeScript
%   itself developed \texttt{dts-gen}, a tool that generates a
%   declaration file that is meant to be used as a \emph{starting point
%     for writing a high-quality declaration file} \cite{dts-gen}.
% \end{description}

Our tool \texttt{dts-generate} is a first step to explore the possibilities for
generating useful declaration files from code
examples. \texttt{dts-generate} comes with a framework that 
supports the generation of declaration files for an existing
JavaScript library published to the \NPM{} registry. The tool gathers
data flow and type information at run time to generate a declaration
file based on that information.

The novelty of our tool is twofold:
\begin{enumerate}
\item
  %% without the heavy lifting of static analysis
  We do not rely on static analysis, which is hard to implement
  soundly and precisely and which is prone to maintenance problems
  when keeping up with JavaScript's frequent language updates.
\item
  Instead we extract example code from the programmer's library
  documentation and rely on dynamic analysis to extract typed usage
  patterns for the library from the example runs.
\end{enumerate}
% The architecture supports the future incorporation of a Symbolic
% Execution Engine that expands the initial code base using the
% signatures in the declaration file. The iterative process of exploring
% new execution paths will refine the generated declaration file in each
% iteration.

@@

\subsection{Contributions}
\label{sec:contributions}

\begin{itemize}
% \item We introduce an architecture that supports generating TypeScript
%   Declaration Files for a given JavaScript Library using run-time
%   information. The architecture supports the future incorporation of a
%   Symbolic Execution Engine that expands the initial code base using
%   the signatures in the declaration file. The iterative process of
%   exploring new execution paths will refine the generated declaration
%   file in each iteration.
\item A framework that extracts code examples from the
  documentation of an \NPM{} package and collects run-time type
  information from running these examples.

\item Design and implementation of the tool \texttt{dts-generate}, a command line
  application that generates a valid TypeScript declaration file for a
  specific \NPM{} package using run-time information.

@@

\section{Motivating Example}
\label{sec:motivating-example}
The \NPM{} package \texttt{glob-to-regexp} is a simple JavaScript library, which turns a glob expression into a regular expression\footnote{https://www.npmjs.com/package/glob-to-regexp}. It
has about 9,900,000 weekly downloads and 188 \NPM{} packages depend on it. If
a developer creates or extends TypeScript code that depends on the
\texttt{glob-to-regexp} library, the TypeScript
compiler and IDE requires a declaration file for that library to
perform static checking and code completion, respectively. With
\texttt{dts-generate} we automatically generate a TypeScript
declaration file for \texttt{glob-to-regexp}. The tool downloads the \NPM{} 
package, runs the examples extracted from its documentation, gathers
run-time information, and generates a TypeScript declaration
file. The result is 
shown in \figref{fig:motivating-example-glob-to-regexp-vscode} and it is
ready for use in a TypeScript project. For example,  Visual Studio's
code completion runs properly with the generated file
(\figref{fig:motivating-example-glob-to-regexp-vscode}). If the
\texttt{glob-to-regexp} package gets modified in the future, a new declaration
file can be generated automatically using
\texttt{dts-generate}. Our comparator tool can compare the new file
for incompatibilities with the previous declaration file.

\begin{figure}[tp]
  \centering
%  \begin{subfigure}{0.70\linewidth}
    \begin{lstlisting}[language=TypeScript,numbers=none]
export = GlobToRegexp;

declare function GlobToRegexp(glob: string, opts ?: GlobToRegexp.I__opts): RegExp;
declare namespace GlobToRegexp {
  export interface I__opts {
    'extended' ?: boolean;
    'globstar' ?: boolean;
    'flags'    ?: string;
  }
}
    \end{lstlisting}
  % \end{subfigure}
  \begin{center}
    \includegraphics[width=0.8\linewidth]{motivating-example-glob-to-regexp-vscode.png}
  \end{center}

  \caption{\textbf{Declaration file for \texttt{glob-to-regexp}
      generated with \texttt{dts-generate}} - The interface is
    detected correctly. Optional parameters are detected. The
    declaration file is usable in Visual Studio
    Code\footnote{https://code.visualstudio.com}.} 
  \label{fig:motivating-example-glob-to-regexp-vscode}
\end{figure}

% After filling in some background information on declaration files,
% Section~\ref{sec:gener-typescr-decl} examines each step in the
% generation process in detail and refer back to this example for
% concreteness.

@@

The declaration file shown in \figref{fig:motivating-example-glob-to-regexp-vscode}
describes a package with a single exported function. The first parameter is of type
\texttt{string} and the second one is an optional interface. TypeScript namespaces
organize types declared in a declaration file and avoid name clashes with other types
\cite{typescript-namespaces}. For example, the interface \texttt{I\_\_opts} declared in
\figref{fig:motivating-example-glob-to-regexp-vscode} is used as
\lstinline[language=TypeScript]{globToRegexp.I__opts} because it is declared within a
namespace.

This file is an instance of one of the standard templates for writing
declaration
files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/templates.html}:
\textbf{module}, \textbf{module-class}, and
\textbf{module-function}. 
Each template corresponds to a different way of describing the exports
of a JavaScript library. The choice of the template depends on the
organization of the underlying JavaScript library:
\begin{description}
\item[module] several exported functions,
\item[module-class] a class-like structure,
\item[module-function] exactly one exported function.
\end{description}

@@

\begin{figure}[tp]
  \centering
  \begin{subfigure}{0.48\linewidth}
    \begin{lstlisting}[language=JavaScript,numbers=none]
var greet = require("./greet-settings-module");

@@

greet({
  greeting: "hello world",
  color: "#00ff00"
});
    \end{lstlisting}
    \caption{Example for \texttt{module-function} template}
    \label{fig:example-module-function}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.48\linewidth}
    \begin{lstlisting}[language=TypeScript,numbers=none]
export = GreetSettingsModule;

declare function GreetSettingsModule(settings: GreetSettingsModule.I__settings): void;
declare namespace GreetSettingsModule {
  export interface I__settings {
    'greeting': string;
    'duration' ?: number;
    'color' ?: string;
  }

@@

  \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=JavaScript,numbers=none]
var myLib = require("./greet-module");

@@

var goodbye = myLib.makeGoodBye();
console.log("The computed goodbye is: " + goodbye);    
      \end{lstlisting}
    \caption{Example for \texttt{module} template}
    \label{fig:example-module}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript,numbers=none]
export function makeGreeting(str: string): string;
export function makeGoodBye(): string;        
      \end{lstlisting}
      \caption{Declaration file for \texttt{module} template}
    \end{subfigure}

    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=JavaScript,numbers=none]
var Greeter = require("./greet-classes-module.js");

var myGreeter = new Greeter("hello, world");
myGreeter.greeting = "howdy";
myGreeter.showGreeting();
      \end{lstlisting}
      \caption{Example for \texttt{module-class} template}
      \label{fig:example-class}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.48\linewidth}
      \begin{lstlisting}[language=TypeScript,numbers=none]
export = Greeter;

@@

  \caption{Example uses and declaration files for different templates}
  \label{fig:typescript-templates-by-example}
\end{figure}

TypeScript provides a guide on how to write high-quality declaration
files\footnote{https://www.typescriptlang.org/docs/handbook/declaration-files/by-example.html}. The guide
explains the main concepts through examples. We selected one example for each template
from the guide and
used \texttt{dts-generate} to generate a corresponding declaration file, as shown in
\figref{fig:typescript-templates-by-example}.

Like \texttt{dts-gen}, a TypeScript package that generates a template for a declaration
file, we determine the kind of declaration file by examining the provided example. If the
imported entity is used solely as a function as in Figure~\ref{fig:example-module-function}, the
generated declaration file follows the module-function template.
If the example only accesses properties of the imported entity as in
Figure~\ref{fig:example-module}, we generate according to the module template.
If the imported entity is used with \lstinline/new/ to create new instances as in
Figure~\ref{fig:example-class}, we generate a module-class template.

The libraries \texttt{glob-to-regexp} as well as \texttt{abs} are both instances of a
\textbf{module-function}.
Most libraries fall in one of these three categories. 
While it is possible to create a library that can be represented with a different template
depending on how it is used, our selection is driven by the examples in the documentation and
hence do reflect the developers intent.
% Run-time analysis proves useful in this case,
% since it is not sufficient to statically analyze the exported module.

% \begin{figure}[tp]
%   \centering
%   \begin{subfigure}{0.80\linewidth}
%       \begin{lstlisting}[language=JavaScript]
% var multiPurposeModule = require('./multi-purpose-module');

% // module-function
% console.log(multiPurposeModule("John", "Doe"));
% // John Doe

% // module-class
% var m = new multiPurposeModule("Jane", "Doe");
% console.log(m.firstName + " - " + m.lastName);
% // Jane - Doe

% console.log(m.sayHello());
% // Hello, my name is Jane Doe

% // module
% console.log(multiPurposeModule.anotherFunction());
% // I am another function! 
%       \end{lstlisting}
%       \caption{index.js}
%     \end{subfigure}

%     \begin{subfigure}{0.80\linewidth}
%       \begin{lstlisting}[language=TypeScript]
% var multiPurposeModule = function(firstName, lastName) {
%   this.firstName = firstName;
%   this.lastName = lastName;

%   return firstName + " " + lastName;
% };

% multiPurposeModule.prototype.sayHello = function() {
%   return "Hello, my name is " + this.firstName + " " + this.lastName;
% };

% multiPurposeModule.anotherFunction = function() {
%   return "I am another function!";
% };

% module.exports = multiPurposeModule;   
%       \end{lstlisting}
%       \caption{multi-purpose-module/index.js}
%     \end{subfigure}
%   \caption{Example module that implements all three templates simultaneously}
%   \label{fig:multi-purpose-module-example}
% \end{figure}

@@

% \begin{figure}[tp]
%   \centering
%   \begin{subfigure}{0.48\linewidth}
%     \begin{lstlisting}[language=JavaScript]
% var multiPurposeModule = require('./multi-purpose-module');

% multiPurposeModule("John", "Doe");
% // John Doe
%     \end{lstlisting}
%     \caption{Example module \texttt{multi-purpose-module} used as \texttt{module-function}}
%   \end{subfigure}
%   \hfill
%   \begin{subfigure}{0.48\linewidth}
%     \begin{lstlisting}[language=TypeScript]
% export = MultiPurposeModule;

% declare function MultiPurposeModule(firstName: string, lastName: string): string;
%     \end{lstlisting}
%     \caption{Generated declaration file with \texttt{module-function} template}
%   \end{subfigure}

%   \begin{subfigure}{0.48\linewidth}
%     \begin{lstlisting}[language=JavaScript]
% var multiPurposeModule = require('./multi-purpose-module');

% var m = new multiPurposeModule("Jane", "Doe");
% var fullName = m.firstName + " - " + m.lastName;
% // Jane - Doe

% m.sayHello();
% // Hello, my name is Jane Doe
%     \end{lstlisting}
%     \caption{Example module \texttt{multi-purpose-module} used as \texttt{module-class}}
%     \end{subfigure}
%     \hfill
%     \begin{subfigure}{0.48\linewidth}
%       \begin{lstlisting}[language=TypeScript]
% export = MultiPurposeModule;

% declare class MultiPurposeModule {
%   constructor(firstName: string, lastName: string);
%   sayHello(): string;
% }

% declare namespace MultiPurposeModule {
% }
%       \end{lstlisting}
%       \caption{Generated declaration file with \texttt{module-class} template}
%     \end{subfigure}

%     \begin{subfigure}{0.48\linewidth}
%       \begin{lstlisting}[language=JavaScript]
% var multiPurposeModule = require('./multi-purpose-module');

% multiPurposeModule.anotherFunction();
% // I am another function!
%       \end{lstlisting}
%       \caption{Example module \texttt{multi-purpose-module} used as \texttt{module}}
%     \end{subfigure}
%     \hfill
%     \begin{subfigure}{0.48\linewidth}
%       \begin{lstlisting}[language=TypeScript]
% export function anotherFunction(): string;
%       \end{lstlisting}
%       \caption{Generated declaration file with \texttt{module} template}
%     \end{subfigure}
